Formalizing the Edmonds-Karp Algorithm

### 2016-03-01 paper/root.tex
modfied first section

\section{Short Background on MinCutMaxFlow and FoFu}
% \begin{comment}
%   GOAL: 
%     1) Remind the reader of FoFu. The educated reader should have the feeling:
%       Yes, now I recall FoFu, and understand (again) what it does.
%       
%     2) Set the field for further, more detailed descriptions in next section.
%   
% 
%   Short description of Network: Finite graph over nodes V, 
%     edge (u,v) annotated with capacity c(u,v).
%     Assuming distinct nodes s and t. 
%     Note on additional assumptions: 
%         No parallel edges, s only outgoing, t only incoming, all nodes on path from s to t. 
%         Can transform any Network to match these assumptions, preserving the flow \cite{???}.
%         
%   s-t Flow: Annotation of edges with values, such that: 
%     1) Values smaller than capacities.
%     2) Kirchhoff law: sum of incoming flows = sum of outgoing flows, for all nodes except s t
%   
%   Value of flow: Incoming flow - outgoing flow of s
%   
%   Max-Flow problem. 
%     Rpt. why important?
%   
%   Cuts, intuition of min-cut >= max-flow
%   
%   Theorem: Min-Cut = Max-Flow
%   Proven via augmenting flow in residual graph. 
%     Present 1,2,3 of min-cut max-flow equivalences
%   
%   This yields an algorithm to compute max-flows, if we can compute flows in residual graph.
%     --> Simple way: Augmenting flow via augmenting path. ==> FoFu-scheme. 
%       Termination? In General: Only for integer capacities.
%       Edmonds-Karp: Shortest augmenting path. Always terminates! Complexity: O(|E||V|) outer loop iterations, BFS requires O|E|.
%   \end{comment}
A flow network is a weighted digraph where the set of vertices (nodes) $V$ is finite, and each edge $(u, v)$ has a real capacity $c (u, v) \geq 0$. There are two distinct vertices $s$ (source) and $t$ (sink) in the flow network. In this case the network is called an $s$-$t$ network. To make the formalization simple, some additional assumptions are added to the definition as following: 1) The source only has outgoing edges while the sink only has incoming edges; 2) If the flow network contains an edge $(u, v)$ then there is no edge $(v, u)$ in the reverse direction; and 3) every vertex of the flow network must be on a path from $s$ to $t$. Although these assumptions seems restrictive, any network may be transformed such that it satisfies all the aforementioned properties.

@@

Assume a flow $f$ in an $s$-$t$ network with capacity function $c$. An edge $(u, v)$ of the network can admit an additional amount of flow equal to the edge’s capacity minus the flow along that edge. If this value is positive, then a $(u, v)$ edge is added to the residual graph with a residual capacity $c_f (u, v) = c (u, v) - f (u, v)$. So, the residual graph only contains those edges of the original flow network that have left-over capacities. The residual graph may also contain some edges that do not exist in the original flow network. For each edge $(u, v)$ in the flow network which has a positive flow value, a $(v, u)$ edge is added to the residual graph with the residual capacity $c_f (v, u) = f (u, v)$. The edge $(v, u)$ indicates the possibility for pushing the flow backward along the edge $(u, v)$ in the original flow network.

After each update of the flow $f$, the algorithm considers the new residual graph $G_f$ corresponding to that flow. Then the algorithm looks for an augmenting path--- a simple path from source to the sink--- in the residual graph. If such a path could be found, it will be used to construct a flow in the residual graph. This flow provides a road map for modifying the current flow in the flow network. The process of increasing the current flow using an augmenting flow is called augmentation. Let $f$ be the current flow in an $s$-$t$ network $G = (V, E)$ with capacity function $c$, and $f$ be a flow in the corresponding residual graph. The augmentation of $f$ using $f'$ is a function from $V \times V$ to $\mathbb{R}$, defined as following:
\[ (f \uparrow f') (u, v) = 
	\begin{cases}
	f (u, v) + f'(u, v) - f'(v, u) \hfill & \text{ if $(u, v) \in E$} \\	
	0 \hfill & \text{ otherwise} \\
	\end{cases} 
\]

@@

\section{Formalizing MinCutMaxFlow and Fofu}
%  GOAL: Present highlights of our formalization, persuade the reader that we have done something substantial.
%    Reader should think: Yeah, they did cool stuff!
%
%  [???* Thematize definition of flows: On Graphs vs. on Networks. You find both in the literature. 
%      We decides for [...]. This is better suited, as is gives nice and elegant proofs, even formally, bla bla bla]
%    
%  Proof of MinCut-MaxFlow: Follows the textbook proof.
%    Uses Isar to even look like a textbook proof, being comprehensible even without running
%      Isabelle/HOL.
%      
%    For example, \isai{augment_flow_presv_cap}.
%      Present formal proof text. Perhaps oppose it to textbook proof?.
%    
%  Abstract Algo looks like pseudocode presented in texbooks.
%    Oppose our algo to textbook pseudocode.
%

>>>>

\section{A short background on the Ford-Fulkerson method}
% \begin{comment}
%   GOAL: 
%     1) Remind the reader of FoFu. The educated reader should have the feeling:
%       Yes, now I recall FoFu, and understand (again) what it does.
%       
%     2) Set the field for further, more detailed descriptions in next section.
%   
% 
%   Short description of Network: Finite graph over nodes V, 
%     edge (u,v) annotated with capacity c(u,v).
%     Assuming distinct nodes s and t. 
%     Note on additional assumptions: 
%         No parallel edges, s only outgoing, t only incoming, all nodes on path from s to t. 
%         Can transform any Network to match these assumptions, preserving the flow \cite{???}.
%         
%   s-t Flow: Annotation of edges with values, such that: 
%     1) Values smaller than capacities.
%     2) Kirchhoff law: sum of incoming flows = sum of outgoing flows, for all nodes except s t
%   
%   Value of flow: Incoming flow - outgoing flow of s
%   
%   Max-Flow problem. 
%     Rpt. why important?
%   
%   Cuts, intuition of min-cut >= max-flow
%   
%   Theorem: Min-Cut = Max-Flow
%   Proven via augmenting flow in residual graph. 
%     Present 1,2,3 of min-cut max-flow equivalences
%   
%   This yields an algorithm to compute max-flows, if we can compute flows in residual graph.
%     --> Simple way: Augmenting flow via augmenting path. ==> FoFu-scheme. 
%       Termination? In General: Only for integer capacities.
%       Edmonds-Karp: Shortest augmenting path. Always terminates! Complexity: O(|E||V|) outer loop iterations, BFS requires O|E|.
%   \end{comment}
A flow network is a weighted digraph where the set of vertices (nodes) $V$ is finite, and each edge $(u, v)$ has a real capacity $c (u, v) \geq 0$. There are two distinct vertices $s$ (source) and $t$ (sink) in the flow network. In this case the network is called an $s$-$t$ network. To make the formalization simple, some additional assumptions are added to the definition as following: 1) The source only has outgoing edges while the sink only has incoming edges; 2) If the flow network contains an edge $(u, v)$ then there is no edge $(v, u)$ in the reverse direction; and 3) every vertex of the flow network must be on a path from $s$ to $t$. Although these assumptions seems restrictive, any network may be transformed such that it satisfies all the aforementioned properties.

@@

%Assume a flow $f$ in an $s$-$t$ network with capacity function $c$. An edge $(u, v)$ of the network can admit an additional amount of flow equal to the edge’s capacity minus the flow along that edge. If this value is positive, then a $(u, v)$ edge is added to the residual graph with a residual capacity $c_f (u, v) = c (u, v) - f (u, v)$. So, the residual graph only contains those edges of the original flow network that have left-over capacities. The residual graph may also contain some edges that do not exist in the original flow network. For each edge $(u, v)$ in the flow network which has a positive flow value, a $(v, u)$ edge is added to the residual graph with the residual capacity $c_f (v, u) = f (u, v)$. The edge $(v, u)$ indicates the possibility for pushing the flow backward along the edge $(u, v)$ in the original flow network.

There are two kinds of edges in the residual graph: 1) those of the residual graph that present the edges of the flow network with left-over capacities; and 2) those of the residual graph that do not occur in the flow network, and indicate the possibility for pushing the flow backward. Formally speaking, consider a flow $f$ on an $s$-$t$ network $G = (V, E)$ with the capacity function $c$. A residual capacity function $c_f \colon V \times V \rightarrow \mathbb{R}$ is defined as following:
\[ c_f (u, v) = 
	\begin{cases}
	c (u, v) - f(u, v) \hfill & \text{ if $(u, v) \in E$} \\
	f (v, u) \hfill & \text{ if $(v, u) \in E$} \\	
	0 \hfill & \text{ otherwise} \\
	\end{cases} 
\]

Then the residual graph corresponding to $f$ and $G$ is indicated by $G_f = (V, E_f)$, where the residual edge set $E_f$ is defined as following:
\[ E_f = \{(u,v) \in V \times V \colon c_f (u, v) > 0\}\]

After each flow update, the algorithm considers the residual graph corresponding to the new flow. Then the algorithm looks for an augmenting path--- a simple path from source to the sink--- in the residual graph. If such a path could be found, it will be used to construct a flow in the residual graph. Let $p$ be such a path. The residual capacity of $p$ is defined as $c_f(p) = min \{c_f(u, v): \text{$(u, v)$ is on  $p$}\}$. Then an augmenting flow $f_p \colon V \times V \rightarrow \mathbb{N}$  in the residual graph can be defined using $p$ as following:
\[ f_p (u, v) = 
	\begin{cases}
	c_f(p) \hfill & \text{ if $(u, v)$ is on $p$} \\	
	0 \hfill & \text{ otherwise} \\
	\end{cases} 
\]

This flow provides a road map for modifying the current flow in the flow network. The process of increasing the current flow using an augmenting flow is called augmentation. Let $f$ be the current flow in an $s$-$t$ network $G = (V, E)$ with capacity function $c$, and $f'$ be a flow in the corresponding residual graph. The augmentation of $f$ using $f'$ is a function from $V \times V$ to $\mathbb{R}$, defined as following:
\[ (f \uparrow f') (u, v) = 
	\begin{cases}
	f (u, v) + f'(u, v) - f'(v, u) \hfill & \text{ if $(u, v) \in E$} \\	
	0 \hfill & \text{ otherwise} \\
	\end{cases} 
\]

@@

\section{Formalizing the Ford-Fulkerson method}
%  GOAL: Present highlights of our formalization, persuade the reader that we have done something substantial.
%    Reader should think: Yeah, they did cool stuff!
%
%  [???* Thematize definition of flows: On Graphs vs. on Networks. You find both in the literature. 
%      We decides for [...]. This is better suited, as is gives nice and elegant proofs, even formally, bla bla bla]
%    
%  Proof of MinCut-MaxFlow: Follows the textbook proof.
%    Uses Isar to even look like a textbook proof, being comprehensible even without running
%      Isabelle/HOL.
%      
%    For example, \isai{augment_flow_presv_cap}.
%      Present formal proof text. Perhaps oppose it to textbook proof?.
%    
%  Abstract Algo looks like pseudocode presented in texbooks.
%    Oppose our algo to textbook pseudocode.
%
